Visualizing demographic evolution using geographically inconsistent census data

### 2019-06-17 paper/intro.tex
GEAN SI submitted version

\section{Intuition}
\label{sec:intuition}
While utterly simple, the network model breaks from the deeply rooted
traditional tabular paradigm in a significant way. Instead of requiring the data
as a collection of fixed entities which properties evolve over time, literally
rows in a table with temporal values as columns, it represents each measurement
as a separate entity and encodes the evolution of these entities over time.

To ease this cognitive transition, we start with an intuitive example of how it
works, using a small portion of a fictitious urban region illustrated on the
left part of Figure~\ref{fig:intuition}. This example includes three different
times ($t_0,t_1,t_2$), with different aggregation areas identified as letters
from A to H. For $t_0$, the initial time, we have areas A and B, with small
houses and a park, respectively. The park remains stable (B, E, and H), but the
houses are partially replaced by larger buildings (C and F).

This example also illustrates some of the challenges of harmonisation. None of
these aggregation areas is clearly suitable as an interpolation target. In fact,
adopting any of these areas as a target would require merging heterogeneous
regions and/or dividing homogeneous regions. For instance, by choosing the
regions of $t_1$, A would be split to match regions C and D, which appears to be
a rather reasonable approximation in this homogeneous artificial example, albeit
the fallacy of division. Region F would be similarly split to match C, but F
would be split and merged with G to match D, potentially leading to statistical
measurements that do not properly represent either region.

@@

%I'm adding this because Jeff Allen (Farber's guy) struggled with this.
Note that the edges of this network merely encode that two regions are related.
This is a binary information, there is no apportionment, no areal measurements,
no population percentages associated with the edge. Indeed, our method also
connects regions of the same time that share borders, representing exactly that
they are neighbouring areas.

>>>>

\section{Intuition}
\label{sec:intuition}
While utterly simple, the network model breaks from the deeply rooted
traditional tabular paradigm in a significant way. The traditional method requires the data to be treated as a collection of fixed entities with properties that evolve over time -- rows in a table with temporal values as columns. By contrast, our method represents each measurement as a separate entity and encodes the evolution of these entities over time.

To ease the cognitive transition to a new paradigm, we start with an intuitive example of how the method
works, using a small portion of a fictitious urban region illustrated on the
left part of Figure~\ref{fig:intuition}. This example includes three different
times ($t_0,t_1,t_2$), with different aggregation areas identified as letters
from A to H. For $t_0$, the initial time, we have areas A and B, with small
houses and a park, respectively. The park remains stable (B, E, and H), but the
houses are partially replaced by larger buildings (C and F).

This example illustrates some of the challenges of harmonisation. None of
these aggregation areas is clearly suitable as an interpolation target. In fact,
adopting any of these areas as a target would require merging heterogeneous
regions and/or dividing homogeneous regions. For instance, by choosing the
regions of $t_1$, A would be split to match regions C and D, which appears to be
a rather reasonable approximation in this homogeneous artificial example (even if it commits the fallacy of division). Region F would be similarly split to match C, but F
would be split and merged with G to match D, potentially leading to statistical
measurements that do not properly represent either region.

@@

%I'm adding this because Jeff Allen (Farber's guy) struggled with this.
Note that the edges of this network merely encode that two regions are related.
This is binary information, there is no apportionment, no areal measurements,
no population percentages associated with the edge. Indeed, our method also
connects regions of the same time that share borders, representing exactly that
they are neighbouring areas.

In the following, we argue that this network representation allows us to study neighbourhood change in a way that does not require an prior interpolation.

### 2019-06-17 paper/study.tex
GEAN SI submitted version

We consider a region that is approximately the administrative border of the
current city of Toronto, using all seven available aspects with equal weights.
While Chicago was fairly stable, Toronto is known to be a more dynamic and
diverse city, with significant and increasing immigrant
population~\citep{hulchanski2007three}. Toronto is also known for a stable and
well defined Jewish community~\citep{Shahar2016}. Therefore, we expect the
combination of stable and dynamic regions on the results, with Place of Birth,
Home Language, and Religion identified as relevant aspects. The results are
summarised in Figure~\ref{fig:to}, considering eight clusters.

@@

The geographical borders of the clusters obtained using our method are similar
to the regions presented by previous studies considering
Toronto~\citep{hulchanski2007three}, but our interface provides a deeper insight
into their demographic composition, since we consider more data than solely
Average Income, which appears to be a good proxy variable nonetheless. This
scenario showcases the ability of our method and interface to capture and
understand the sources of urban volatility.

@@

We selected for this scenario a region around the metropolitan area of Los
Angeles (LA), following urban density. While Chicago and Toronto demonstrated
the abilities of our method for stable and dynamic cities, LA is considerably
larger, both in population and area. Since our method identifies only the eight
most distinct groups, we expect that some groups not to be identified,
especially if their characterisation is similar to another groups. Further, our
system does not include the census variables related to Hispanic heritage, which
were only included in the newer censuses. However, from the literature, we
expect to see some clusters where Race is an important aspect, including White,
Black, and Asian~\citep{Reibel2003}.

@@

These results are visually similar the analysis presented by
Dellmelle~\citep{Delmelle2016}, what was there identified as 'persistently
struggling' corresponds to the purple cluster in Figure~\ref{fig:la}; some
portions of the 'stable elite' cluster, near Inglewood, correspond to the high
income, high education yellow cluster. The other groups identified by that study
were clumped by our system into the green cluster, which would be further
divided if more than eight clusters were to be used.

>>>>

We consider a region that is approximately the administrative border of the
current city of Toronto, using all seven available aspects with equal weights.
While Chicago was fairly stable, Toronto is known to be a more dynamic and
diverse city, with significant and increasing immigrant
population~\citep{hulchanski2007three,Fong2011}, especially Asian~\citep{Fong2003}. 
Toronto is also known for a stable and well defined Jewish community~\citep{Harold2018, Fong2011}. Therefore, we expect the combination of stable and dynamic regions on the results, with Place of Birth,
Home Language, and Religion identified as relevant aspects. The results are
summarised in Figure~\ref{fig:to}, considering eight clusters.

@@

The geographical borders of the clusters obtained using our method are similar
to the regions presented by previous studies considering
Toronto~\citep{hulchanski2007three}. However, our interface provides a deeper insight
into their demographic composition, since we consider more data than solely
Average Income, which appears to be a good proxy variable nonetheless. This
scenario showcases the ability of our method and interface to capture and
understand the sources of urban volatility.

@@

We selected for this scenario a region around the metropolitan area of Los
Angeles (LA), following urban density. While Chicago and Toronto demonstrated
the abilities of our method to examine stable and dynamic cities, LA is considerably
larger, both in population and area. Since our method identifies only the eight
most distinct groups, we expect that some groups will not be identified,
especially if their characterisation is similar to another groups. Further, our
system does not include the census variables related to Hispanic heritage, which
were only included in the newer censuses. However, from the literature, we
expect to see some clusters where Race is an important aspect, including White,
Black, and Asian~\citep{Reibel2003}.

@@

These results are visually similar to the analysis presented by
Dellmelle~\citep{Delmelle2016}.  What was there identified as 'persistently
struggling' corresponds to the purple cluster in Figure~\ref{fig:la}; some
portions of the 'stable elite' cluster, near Inglewood, correspond to the high
income, high education yellow cluster. The other groups identified by that study
were clumped by our system into the green cluster, which would be further
divided if more than eight clusters were to be used.
